\textbf{\sc Towards Deliberative Control of Autonomous Robots}

### 2012-01-31 arch.tex
updated deliberation section of trex to refer to solve algorithm

The overal execution cycle of the agent is emerging through these
interaction and the processing that occurs in each reactors in order
to process them. New observation occurring at every tick inform each
reactor of the current state of the world and impose them to identify
how it impact their state (and potentially what they planned for the
future). Goals they receive will trigger in the reactor the need to
identify a strategy (or plan) that will allow them to reach these new
states knowing their current state deduced from synchronization.When
such a plan has been identified, the {\em external}  part is then
dispatched to other reactors that can do the same. This flow and focus
around state and goals is perfectly fitting with automated planning
and even more so to the Europa framework. As an illustration, in 
 Fig. \ref{fig:agent}, only \textsf{Shore comm.} and \textsf{Vehicle}
 reactors are specialized C++ code (acting as interface to external
 programs). All the other reactors are based on the europa framework
 differentiated only by the domain model they manipulate and som
 heuristic configuration parameters. In the following sections we are
 going to focus specifically ion this kind of reactor to illustrate
 how  a formal planning framework can be used in our architecture to
 both track state evolutions and plan for future objectives while
 remaining efficient enough to do these tasks in a real-time control 
loop.

@@

The planning process inside a reactor is directly based on the way
europa work which was described in previous section. Still some simple
alterations were made in  order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in 
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with 
the deliberation of other reactors\footnote{Current implementation of
  T-REX is running on a single process and it is the responsibility of
  the agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason we 
decompose the algorithm \ref{alg:europa:solve} inside the reactor by
having the recursive calls decomposed in single steps. This step
function will be called after synchronization and as long as the
previous call in this tick did not return \textsf{true} -- which would
express that the current solution is complete.
{\em will need to change that in algorithm environment and refine it}

\begin{verbatim}
bool step(PartialPlan plan, DecisionStack stack) 
   if( isInconsistent(plan) ) {
     relaxPlan(plan, stack);
     notify("No plan");
     return true;
   } 
   Flaw flaw = chooseFlaw(flaw);
   if( flaw==NULL )
     return true;
   DecisionPoint decision = makeDecisionPoint(flaw, plan);
   while(1) {
     while( decision.hasNext() ) {
       PartialPlan pp = decision.executeNext();
       if( isInconsistent(pp) ) {
         decision.undo();
       }  else {
          plan = pp;
          stack.push(decision);
          return chooseFlaw(flaw)==NULL;
       }
     }
     if( decisionStack.empty() ) {
       notify("No plan");
       return true;
     } else {
        repeat {
          decision = decisionStack.pop();
          decision.undo();
        } until( decision.hasNext() );
        return false;
     }
   } 
}
\end{verbatim}

@@

When the number of flaw is 0 A plan solution is found. This has two
effects:
\begin{enumerate}
\item This reactor does not need to deliberate until the next synchronization.
\item The {\em external} part of this plan can then be posted which
  eventually (at the end of the current tick if these goals overlap
  the owner planning window) will be dispatched as goals triggering
  new deliberation on these reactors.
\end{enumerate}

>>>>

The choice of representation of state variables, observations and
goals is directly derieved to the timleines and tokens from europa. 
such representation is ich enough to allow the excahnge of state
information in a formal and yet flexible way and its transaltion in
planning frameworks that have an explicit representation of time 
(such as but not limited to europa) is often trivial.

The overall execution cycle of the agent is also abstracted as some
continuous planning/deliberation cycle between all the
reactors. Indeed , at the agent interface level, each reactor provided
only two abstract calls that conceptualize their execution:
\begin{itemize}
\item \texttt{synchronize} that takes the last {\em external}
  observations as argument and will return the {\em internal}
  observations for this reactor or a failure report.
\item \texttt{step} that accept new goals requested (if any) on the
  {\em internal} timeline of this reactor and will execute one step of
  deliberation. The returned value indicates if a complete plan has
  been found and in that case will also provide the part of this plan
  applied to {\em external} timelines of this reactor.
\end{itemize}
Both of these two deliberative functions have different scope and
temporal constraints. The \texttt{synchronize} call is considered as
atomic and will focus only on state inference from the current tick by
identifying the current value of the {\em internal} state variables
from the latest updates on the {\em external} variables. This process
will then propagate up in the reactor graph at any single tick to
ensure that all reactors are aware of the current state of the
world. \texttt{step} relates to a single decision step of the planning
process which is focused on the future evolution of the state
variables and it can take multiple steps spanning other multiple
ticks for a reactor to produce a plan that is both complete and
valid. when such plan is identified its {\em external} part is then
dispatched to the corresponding reactors which in turn can start their
deliberation \texttt{step}s in order to satisfy these new
objectives. As a consequence physical execution will occur as these
goals and planning propagate down in the reactor graph until they
eventually become goals for the lowest level reactor which -- instead
of deliberating -- may just transform this goal in a physical command
execution.

The way the agent interleave planning \texttt{step}s and
synchronization is illustrated in Figure \ref{fig:tick-exec}.
Synchronization happens at every single tick and will interrupt
planning allowing this reactor to identify its current state that can
propagate through the reactor graph in order to ensure t aconsitent
view of the state of the world between all the reactors.  
\begin{figure}[!htbp]
  \centering
  \vskip-1pc
  \includegraphics[width=0.55\columnwidth]{figs/tick-cycle}
  \caption{\small \texttt{T-REX} reactor execution cycle: {\em
      Deliberation} is interrupted by {\em synchronization} at the
    beginning of every {\em tick} allowing integration of state
    information.}
  \label{fig:tick-exec}
  \vskip-0.8pc
\end{figure}
This view of the reactor has two parallel tasks aloow for long
planning deliberation while ensuring that the state evolution
propagates through the agent as time advance. These two tasks 
intertwined together give the two natural flows (bottom-up for state
estimation and top to bottom for future plan) that we expect to see in
a control loop. As they are both conceptualized as inference processes
it is natural to implement a reactor based on generic automated
planning. For example in Fig. \ref{fig:agent}, appart \textsf{Shore
  comm.} and \textsf{Vehicle} acting as interfaces to the system,
every other reactors (colored in blue) are different instance of our
europa reactor differntiated only by their model and europa solver
configurations.

@@

The planning process inside a reactor is directly based on the way
europa work which was described in previous section. Still some simple
alterations were made in  order to both include deliberation
information one reactor provide to the architecture -- such as its
deliberation latency and its planning look-ahead both expressed in 
ticks -- and the fact that we potentially need to interleave this
deliberation with not the synchronization process but also with 
the deliberation of other reactors\footnote{Current implementation of
  T-REX is running on a single process and it is the responsibility of
  the agent itself to emulate the multi-threading of the reactors
  deliberation and synchronization}. For this reason while the overal
planning of a europa reactor strictly relies on the europa framework
plnanning solver, its slices the execution of Algorithm
\ref{alg:europa:solve} in atomic steps. The interruption is done
around the recursive call (line \ref{li:recurse}) while always
ensuring that the current partial plan mainitained is never poroven
inconsitent at any single step. This can be managed by allowing the
call to backtrack until a DecisionPoInt that is not exhausted (ie
$decision.hasNext()$ is \texttt{true}) is find in the call stack.

@@

The scope of planning for the reactor plays also a role in the
deliberation process as we limit deliberation to the sliding temporal
window $[\tau+\lambda, \tau+\lambda+\pi]$ where:
\begin{itemize}
\item $\tau$ is the current tick
\item $\lambda$ is the specified latency of this reactor. Which is an
  indicator on the expected maximum time this reactor is expected
  before being able to produce a plan.\footnote{Do note that this
    parameter remains indicative and, while it is highly recommended to
    select a sound value, a failure to produce a plan in this delay is
    not considered as a critical failure of the reactor}
\item $\pi$ is the planning look-ahead of the reactor. which indicates
  how many ticks in the future this reactor is looking ahead while planning.
\end{itemize}
This scope allow to filter out any tokens that are either necessarily
ending before $\tau$ or necessarily starting after. Focusing the
[planning problem to only the tokens that are in a reasonable
future. This window is taken into account by the agent which will
notify the reactor of new {\em internal} goals only when they overlap
this window but also at the deliberation solver in order to reduce the
number of tokens to be evaluated and consequently reduce the cost of
deliberation. As time advance flaws that where initially beyond this
window will become active flaw that the reactor can then evaluate
resulting on an apparent continuous planning of the reactor.

When there's no more flaw is present in the current planning scope a
plan  solution is found. This has two effects:
\begin{enumerate}
\item This reactor does not need to deliberate until the next synchronization.
\item The {\em external} part of this plan can then be posted which
  eventually (at the end of the current tick if these goals overlap
  the owner planning window) will be dispatched as goals triggering
  new deliberation on these reactors.
\end{enumerate}
